Scheduling and Vectorization for MPC

### 2022-03-15 prob_stmt_np_hardness.tex
first iteration of NP-Hardness proof

\ana{Commenting out this section to have paper compile on Github. Will revert!}
%\begin{comment}

@@

\ana{Commented out this section to have paper compile on Github. Will revert!}

\begin{comment}
\ishaq{The follow text is all from Linsay's thesis}.
\brevised{
We consider two operations, call them $A$ and $M$. $A$ and $M$ are two abstract MPC instruction, 
but as an example, $A$ stands for the ADD MPC instruction 
and $M$ stands for the MUL instruction. Each instruction in the program is either an $A$-instruction or
an $M$-instruction. In order to benefit from parallelization/amortization, we must schedule two or more 
$A$-instructions in the same parallel node (or two or more $M$-instructions in the same parallel node). 
Scheduling $A$-instructions in parallel with $M$-instruction does not benefit from amortization.
It incurs the exact same cost as scheduling the $A$-instructions in a node $P_A$, scheduling the $M$-instructions 
in a node $P_M$, and having $P_A$ precede $P_M$ in the parallel schedule. This is the difference 
between classical scheduling, as studied in parallel computing, and MPC scheduling.

In our theoretical treatment, we make the following assumptions:

\begin{enumerate}

\item $A$ and $M$ are of equal cost, 1 unit.
\item There is unlimited bandwidth---i.e., a single $A$-instruction (or $M$-instruction) costs as much as $N$ amortized $A$-instructions 
(or $M$-instructions), namely 1 unit.

\end{enumerate}

%  Furthermore the run time cost of 
% computing operations of the same type remains the same regardless of the number of operations that are run.
% Roughly speaking, 1 A operation costs the same as 100 A operations and 1000 A operations.

Consider a loop body that consists of $n$ sequences: $S_1$, ... $S_n$ of $A$ and $M$ instructions. 
More precisely, the loop body is such that its instructions can be grouped into such sequences. 
$S_1$, ... $S_n$ can execute in parallel, however, all instructions within a sequence must 
execute sequentially. For example, consider the three sequences (the right arrow indicates a \emph{dependence},
meaning that the source node must execute before the target node): 
\begin{enumerate}
\item $A \rightarrow M \rightarrow A$
\item $A \rightarrow A \rightarrow A$
\item $M \rightarrow A \rightarrow M$
\end{enumerate}

A \emph{schedule} $P: P_1 \rightarrow P_2 \dots \rightarrow P_k$ is such that for each sequence 
$S_i$ in the set, if $S_i[k]$ precedes $S_i[k']$ in $S_i$ then $S_i[k]$ is scheduled in node $P_l$, $S_i[k]$ 
is scheduled in node $P_{l'}$, and $P_l$ precedes $P_{l'}$ in $P$.

@@

where $\mathit{cost}(P_i) = 1$ if $P_i$ consists of $A$-instructions only, or $M$-instructions only, 
and $\mathit{cost}(P_i) = 2$ if $P_i$ mixes $A$-instructions and $M$-instructions.

The problem is to find a schedule $P$ with \emph{minimal cost}. For example, 
a schedule with minimal cost for the sequences above is
\[ A(1), A(2) \rightarrow M(1), A(2), M(3) \rightarrow A(1), A(2), A(3) \rightarrow M(3) \]
(The parentheses above indicate the sequence where the instruction comes from: (1), (2), or (3).)
The cost of this schedule is 5.

The problem of finding a schedule $P$ with a minimal $cost(P)$ for a given loop body has been shown
to be an NP-Hard problem, as it can be reduced to the problem of finding a \emph{shortest common supersequence}, 
a known NP-Hard problem\cite{Maier1978},\cite{Vazirani2010}. The shortest common supersequence problem is as follows: 
given two or more sequences find the the shortest sequence that contains all of the original sequences. This can be solved
in $O(n^k)$ time, where $n$ is the cardinality of the longest sequence and $k$ is the 
number of sequences. For our problem $n$ is the maximum length of a node and $k$ is the number
of total number of nodes. %\ana{I think it is clear without additional explanation...}
%\ana{Also, add the full statement of the problem here. Now it's not clear what's NP-hard, what's n, what's k, etc.?}
%\lindsey{not exactly sure what need to go it there, but does that make it clearer?}

To see the reduction, 
suppose $P$ is a schedule with minimal cost (computed by a black-box algorithm). 
We can derive a schedule $P'$ with the same cost as $P$, by mapping each mixed node $P_i \in P$ 
to two consecutive nodes in $P'$: an $A$-instruction node followed by an $M$-instruction node.
Clearly, $P'$, which now is a sequence of $A$'s and $M$'s, is a supersequence of each sequence 
$S_i$, i.e., $P'$ is a common supersequence 
of $S_1 \dots S_n$. It is also a shortest common supersequence. (To see this, suppose, a 
shorter common supersequence, $P''$, exists. $P''$ is a schedule of $S_1 \dots S_n$
and the $\mathit{cost}(P'')$ equals the length of $P''$. Since $P'$ is longer than
$P''$, and $\mathit{cost}(P') = \mathit{cost}(P)$, that means $\mathit{cost}(P'') < \mathit{cost}(P)$, 
which is a contradiction since $P$ is a schedule with minimal cost.)
}
\end{comment}

>>>>

\ishaq{TODO: make it consistent with the next section.}

@@

\ishaq{TODO: need make amortized cost a function (like in the problem statement above), it is proving to be tricky.}

We consider two operations, call them $A$ and $M$. $A$ and $M$ are two abstract MPC instruction, but as an example, $A$ stands for the ADD MPC instruction and $M$ stands for the MUL instruction. Each instruction in the program is either an $A$-instruction or an $M$-instruction. In order to benefit from parallelization/amortization, we must schedule two or more $A$-instructions in the same parallel node (or two or more $M$-instructions in the same parallel node). We also assume that scheduling $A$-instructions in parallel with $M$-instruction does not benefit from amortization\footnote{this is not strictly true, but assuming it, e.g. as in \cite{Ishaq2019, Demmler2015ABYA, Mohassel2018}, helps with the exposition.}. It incurs the exact same cost as scheduling the $A$-instructions in a node $P_A$, scheduling the $M$-instructions in a node $P_M$, and having $P_A$ precede $P_M$ in the parallel schedule.

We use the following cost model:

\begin{enumerate}
    \item $A$ costs $\alpha$ units and $M$ costs $\beta$ unites.
    \item There is unlimited bandwidth i.e. a single $A$-instruction (or $M$-instruction) costs as much as $N$ amortized $A$-instructions (or $M$-instructions), concretely either $\alpha$ unites or $\beta$ units.
\end{enumerate}

Consider a loop body that consists of $n$ sequences: $S_1$, ... $S_n$ of $A$ and $M$ instructions. More precisely, the loop body is such that its instructions can be grouped into such sequences. $S_1$, ... $S_n$ can execute in parallel, however, all instructions within a sequence must execute sequentially. For example, consider the three sequences (the right arrow indicates a \emph{dependence}, meaning that the source node must execute before the target node):

\begin{enumerate}
    \item $A \rightarrow M \rightarrow A$
    \item $A \rightarrow A \rightarrow A$
    \item $M \rightarrow A \rightarrow M$
\end{enumerate}

A \emph{schedule} $P: P_1 \rightarrow P_2 \dots \rightarrow P_k$ is such that for each sequence $S_i$ in the set, if $S_i[j]$ precedes $S_i[j']$ in $S_i$ then $S_i[j]$ is scheduled in node $P_\ell$, $S_i[j']$ is scheduled in node $P_{\ell'}$, and $P_\ell$ precedes $P_{l\ell'}$ in $P$.

@@

where $\mathit{cost}(P_i) = \alpha$ if $P_i$ consists of $A$-instructions only, $\beta$ if $P_i$ consists of $M$-instructions only, and $\mathit{cost}(P_i) = \alpha + \beta$ if $P_i$ mixes $A$-instructions and $M$-instructions.

The problem is to find a schedule $P$ with \emph{minimal cost}. For example, a schedule with minimal cost for the sequences above is \[ A(1), A(2) \rightarrow M(1), A(2), M(3) \rightarrow A(1), A(2), A(3) \rightarrow M(3) \] (The parentheses above indicate the sequence where the instruction comes from: (1), (2), or (3).) The cost of this schedule is $3\alpha + 2\beta$.

The problem of finding a schedule $P$ with a minimal $cost(P)$ for a given loop body has been shown to be an NP-Hard problem, as it can be reduced to the problem of finding a \emph{shortest common supersequence}, a known NP-Hard problem\cite{Maier1978},\cite{Vazirani2010}. The shortest common supersequence problem is as follows: {\it given two or more sequences find the the shortest sequence that contains all of the original sequences.} This can be solved in $O(n^k)$ time, where $n$ is the cardinality of the longest sequence and $k$ is the number of sequences. For our problem $n$ is the maximum length of a node and $k$ is the number of total number of nodes.

To see the reduction, suppose $P$ is a schedule with minimal cost (computed by a black-box algorithm). We can derive a schedule $P'$ with the same cost as $P$, by mapping each mixed node $P_i \in P$ to two consecutive nodes in $P'$: an $A$-instruction node followed by an $M$-instruction node. Clearly, $P'$, which now is a sequence of $A$'s and $M$'s, is a supersequence of each sequence $S_i$, i.e., $P'$ is a common supersequence of $S_1 \dots S_n$. It is also a shortest common supersequence. To see this, let $X$ and $Y$ denote, respectively, the number of $A$ and $M$ nodes in $P'$. The cost of $P'$, and $P$, is $X \cdot \alpha + Y \cdot \beta$. Now suppose, there exists a shorter common supersequence, $P''$ that consists of $X'$ nodes of type $A$-instructions $Y'$ nodes of type $M$-instructions. Since $P''$ is shorter than $P'$, therefore $X' + Y' < X + Y$, and $X' \cdot \alpha + Y' \cdot \beta < X \cdot \alpha + Y \cdot \beta$ i.e. $\mathit{cost}(P'') < \mathit{cost}(P')$. But $\mathit{cost}(P') = \mathit{cost}(P)$ and $\mathit{cost}(P)$ is the optimal cost.  Therefore $\mathit{cost}(P'') < \mathit{cost}(P')$ is contradiction and no such $P''$ exists.
